\documentclass[../primer.tex]{subfiles}

\begin{document}

\chapter{Formulate}
%% --------------------------------------------------

\section{Data}
%% --------------------------------------------------
Data are our quantitative anchor to reality.\footnote{One could
  \href{http://phdcomics.com/comics.php?f=1816}{make a case} for data in the
  plural (data are facts) or in the singular (data is information). We will use
  the former.} They allow us to make numerical statements about the physical
world, and thus are invaluable to solving problems. However, data are not
infallible -- one must have both skepticism and appreciation of variability in
order to use data effectively. We will consider two kinds of data in this
primer:

\textbf{Physical Data} are comprised of measurements of the physical world. The
vast majority of physical measurements involve the comparison of some physical
quantity of interest against a defined standard; this is most obviously seen in
simple measurements of length, where one can compare a meter stick against a
physical object. More complicated measurements involve some quantity which
cannot be directly measured, but can instead be obtained through a
transformation based on a physical model. An example: It would be challenging to
measure pressure directly. However, we can build a barometer out of a sealed
body with fluid, which essentially converts pressure changes into length
changes. Carrying out the transformation from measured lengths to pressures
requires an appeal to hydrostatics, which carries with it some particular
assumptions.

Variation in physical data arises from \emph{unknown variables}, what we
sometimes call \emph{lurking variables}.\cite{box1966} In the barometer example
above, suppose that our instrument experienced both temperature and pressure
changes. If our barometer used a fluid whose density changed with the pressure,
\emph{and if we did not account for the temperature changes}, then even at a
constant external pressure, we might measure fluctuating values. In this case,
we would call temperature a lurking variable. Detecting and controlling lurking
variables is challenging, but necessary to improve physical
measurements.\cite{joiner1981,delRosario2017lurking}

Luckily, the variations in physical data tend to exhibit `nice' properties, and
so often lend themselves to statistical characterization. For this reason, the
tools of probability and statistics are very well-suited to tackling physical
data.

\textbf{Simulation Data} are the result of models. They are connected to reality
only insofar as their generating model is connected to reality. However,
simulation data has a huge advantage over physical data; one can use simulation
data to make quantitative statements about reality \emph{without physical
  testing}. This can be useful for overcoming constraints of cost (as with
aircraft testing) or legality (as with nuclear weapons). Perhaps one of the
greatest triumphs of simulation is human spaceflight. Using little more than
Newton's laws, early pioneers of spaceflight managed to compute trajectories
that sent astronauts to the moon and back -- physical data alone could not have
stood up this effort.

Of course, we should not oversell the value of simulation. The Apollo program
was built on top of the experience and physical data gathered from Project
Gemini, and before that Mercury. Models themselves are built from studying
physical data, and always carry some assumptions -- assumptions which may be
disconnected from reality. While physical data are subject to variability,
models are often subject to \emph{discrepancy}, often modeled as a difference
between a model prediction and the 'true' value that would arise from a perfect
measurement in physical reality.\cite{higdon2004calibration-prediction}

Despite the fundamental difference between variability and discrepancy,
discrepancy is often modeled using probability and statistics as
well.\cite{kennedy2001bayesian,higdon2004calibration-prediction} As I am writing
this sentence, there is an open debate about the suitability of this approach.
Rather than wade into this argument, we will accept the use of probability for
discrepancy as normative, and proceed to treat both physical and simulation data
with the same toolkit.

\subsection{Summaries}
%% -------------------------

\subsection{Distributions}
%% -------------------------

\section{Models}
%% --------------------------------------------------

\section{Questions}
%% --------------------------------------------------


\end{document}
